---
title: "Reproducible Research - Assignment1"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
Load the libraries required for assignment



```{r cars}

library(knitr)
library(lubridate)
library(ggplot2)
library(dplyr)
```

Read the data, convert the date to date format and omit the observation with NA
```{r}
setwd("C:/Users/vikra/OneDrive/Data Science/Course 5 - Reproducable research/data")
data<-read.csv("activity.csv",header=TRUE, sep = ',',colClasses =c("numeric","character","integer"))
data$date<-ymd(data$date)
datafile<-na.omit(data)
```
Group by date and summarize the steps to plot the histogram
```{r}
datafile.day<-group_by(datafile,date)
datafileday<-summarize(datafile.day,steps=sum(steps))

ggplot(datafileday,aes(steps))+
    geom_histogram(fill = "blue", binwidth = 1000) +
    labs(title = "Histogram of Steps per day", x = "Steps per day", y = "Frequency")

mean(datafileday$steps)
median(datafileday$steps)
```
Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
datafileinterval<-group_by(datafile,interval)
datafileinterval<-summarise(datafileinterval, steps=mean(steps))
ggplot(datafileinterval,aes(interval,steps))+geom_line()
datafileinterval[which.max(datafileinterval$steps),]

data_full <- data
nas <- is.na(data_full$steps)
avg_interval <- tapply(data_full$steps, data_full$interval, mean, na.rm=TRUE, simplify=TRUE)
data_full$steps[nas] <- avg_interval[as.character(data_full$interval[nas])]


steps_full <- data_full
steps_full<-group_by(steps_full,date)
steps_full<-summarize(steps_full, steps=sum(steps))

ggplot(steps_full, aes(x = steps)) +
    geom_histogram(fill = "firebrick", binwidth = 1000) +
    labs(title = "Histogram of Steps per day, including missing values", x = "Steps per day", y = "Frequency")

mean_steps_full <- mean(steps_full$steps, na.rm = TRUE)
median_steps_full <- median(steps_full$steps, na.rm = TRUE)

mean_steps_full
median_steps_full

data_full <- mutate(data_full, weektype = ifelse(weekdays(data_full$date) == "Saturday" | weekdays(data_full$date) == "Sunday", "weekend", "weekday"))
data_full$weektype <- as.factor(data_full$weektype)


interval_full <- data_full %>%
    group_by(interval, weektype) %>%
    summarise(steps = mean(steps))
ggplot(interval_full, aes(x=interval, y=steps, color = weektype)) +
    geom_line() +
    facet_wrap(~weektype, ncol = 1, nrow=2)

```